### 2018-06-18 source/Ch1-Foreword/foreword.tex
why cheesoid exist
added flux-cut catalog

\chapter{Foreword}
\pg
write introduction to the thesis

>>>>

\chapter*{Foreword}

\pg
J.R.R Tolkien wrote, in his Mythopoeia, that ``He sees no stars who does not see them first, of living silver made that sudden burst, to flame like flowers beneath the ancient song"\footnote{See \href{http://home.agh.edu.pl/~evermind/jrrtolkien/mythopoeia.htm}{here} for the full text.}. In his defense of myth-making, he formulates the argument that the attribution of meaning is an act of creation - that ``trees are not `trees' until so named and seen" - and that this capacity for creation defines the human creature. The scientific endeavour is then a social expression of a fundamental feature of humanity, and from this endeavour flows much understanding. This thesis, one thread among many, focuses on the study of astronomical objects as seen by the radio waves they emit.

\pg
What are radio waves? Electromagnetic waves were theorised by James Clerk Maxwell \citep{maxwell} in his great theoretical contribution to modern physics, their speed matching the speed of light as measured by Ole Christensen R\o mer and, later, James Bradley. It was not until Heinrich Rudolf Hertz's 1887 experiment that these waves were measured in a laboratory, leading to the dawn of radio communications - and, later, radio astronomy. The link between radio waves and light was one of association: light is known to behave as a wave (Young double-slit experiment), with the same propagation speed as electromagnetic radiation. Light ``proper" is also known to exist beyond the optical regime: Herschel's experiment shows that when diffracted through a prism, sunlight warms even those parts of a desk which are not observed to be lit (first evidence of infrared light). The link between optical light and unseen electromagnetic radiation is then an easy step, and one confirmed through countless technological applications (e.g. optical fiber to name but one). And as soon as this link is established, a question immediately comes to the mind of the astronomer: what does the sky, our Universe, look like to the radio ``eye"?

\pg
Radio astronomy has a short but storied history: from Karl Jansky's serendipitous observation of the centre of Sagittarius A*, which outshines our Sun in the radio regime, in 1933, to Grote Reber's hand-built back-yard radio antenna in 1937, which successfully detected radio emission from the Milky Way itself\footnote{\url{https://www.nrao.edu/whatisra/hist_reber.shtml}}, to such monumental projects as the Square Kilometer Array and its multiple pathfinders, it has led to countless discoveries and the opening of a truly new window on the Universe. The work presented in this thesis is a contribution to this discipline - the culmination of three years of study, which is a rather short time to get a firm grasp of radio interferometry both in theory and in practice. The need for robust, automated methods - which are improving daily, thanks to the tireless toil of the scientists in the field - is becoming ever stronger as the SKA approaches, looming large on the horizon; but even today, in the precursor era of LOFAR, MeerKAT and other pathfinders, it is keenly felt. When I started my doctorate, the sheer scale of the task at hand felt overwhelming - to actually be able to contribute to its resolution seemed daunting indeed!

\pg
Thankfully, as the saying goes, no society sets for itself material goals which it cannot achieve. This thesis took place at an exciting time for radio interferometry: at the start of my doctorate, the LOFAR international stations were - to my knowledge - only beginning to be used, and even then, only tentatively; MeerKAT had not yet shown its first light; the techniques used throughout my work were still being developed. At the time of writing, great strides have been made. One of the greatest technical challenges of LOFAR - imaging using the international stations - is starting to become reality. This technical challenge is the key problem that this thesis set out to address. While we only achieved partial success so far, it is a testament to the difficulty of the task that it is not yet truly resolved.

\pg
One of the major results of this thesis is a model of a bright resolved source near a famous extragalactic field: properly modeling this source not only allows the use of international LOFAR stations, but also grants deeper access to the extragalactic field itself, which is otherwise polluted by the 3C source's sidelobes. This result was only achieved thanks to the other major result of this thesis: the development of a theoretical framework with which to better understand the effect of calibration errors on images made from interferometric data, and an algorithm to strongly mitigate them.

%
%I owe a debt of gratitude to all my supervisors, whose efforts allowed me to grow (slowly and fitfully) into an adept of radio interferometry. I also owe a debt of gratitude to the European and South African radio-interferometric communities which took me into their fold, for it takes a village to raise a scientist - and I hope to one day be able to pay the favour forwards.

\pg
The structure of this manuscript is as follows: we begin with an introduction to radio interferometry, LOFAR, and the emission mechanisms which dominate for our field of interest. These introductions are primarily intended to give a brief overview of the technical aspects of the data reduced in this thesis. We follow with an overview of the Measurement Equation formalism, which underpins our theoretical work. This is the keystone of this thesis.

\pg
We then show the theoretical work that was developed as part of the research work done during the doctorate - which was published in Astronomy \& Astrophysics. Its practical application - a quality-based weighting scheme - is used throughout our data reduction. This data reduction is the next topic of this thesis: we contextualise the scientific interest of the data we reduce, and explain both the methods and the results we achieve.

### 2018-06-18 source/Ch6-loresEGS/lowres-EGS.tex
added flux-cut catalog

\pg
In this section, we describe the work done to make a wide-field image of the entire LOFAR primary beam, centred on EGS. A large-field, image of this field has never yet been made at such low frequencies. Due to technical limitations, the maximum attainable image size (in terms of number of pixels) will limit resolution to about $5''$. As such, the international stations are not used for this part of the project.
A more complete sky model can be created from this image, and %, allowing for one last round of self-calibration on the 3C295 high-resolution model. 
bright out-of-field sources can be subtracted from high-resolution images of the Extended Groth Strip so that their sidelobe emission do not contaminate the final images. To achieve this, we require direction-dependent calibration. Because we do not use the international LOFAR stations, we do not use our high-resolution model of 3C295. %  This ensures that we do not ``hard-set" errors in our high-resolution model caused by other sources in the field.

@@

%\subsection{Data \& Observation Properties}
%\pg
%The data used to make this image are the same as used to make the high-resolution model of 3C295 in \cref{section.3c295}, but we do not use the visibilities of baselines including international stations in our calibration or imaging. They were acquired through the LOFAR Long-Term Archive service. We do not phase-shift our data when imaging.

%\subsection{Calibration Strategy}

@@

\pg
These sources nevertheless remain in the field, as shown in \cref{image.egs.widefield.ddsc.zoom}. Note that the noise level is not constant throughout the image: this is because the array is not uniformly sensitive to the entire sky, but has a specific beam response. Correcting for this beam response in the image-plane gives integrated flux rather than apparent flux. % gives rise to this effect.

@@

\pg
We select 12 sources in the wide-field image. Our selection is primarily based on whether sources have interesting or peculiar diffuse emission, though some are chosen for being particularly classic examples of radio galaxies. We scan the image from south to north and east to west (scanning upwards then shifting to the right in the image). The source positions are given in \cref{table.egs.sources}.

@@

\pg
However, the scientific value of this image is not limited to its use in imaging the Extended Groth Strip using international LOFAR. Much of the image will not be recreated using the international stations, in part due to computational constraints (long run time for such a large task) and in part due to technical limitations (performing direction-dependent calibration using international LOFAR would likely require the use of techniques as-of-yet unimplemented in the DDF-kMS pipeline). As such, we look in the large-field image for interesting or unusual sources, and attempt to match them to known sources in other bands in order to learn about them. 12 sources were selected, and the most interesting were consistently matched to brightest galaxies in clusters. This indicates that there are many sources with very interesting diffuse emission in and around the Extended Groth Strip, and some of which merit further observations to better resolve their turbulent structure.

@@

\pg
\textcolor{red}{show image of 3C295 calibrated w/ the full source extraction model of the primary beam, compared to result of self-cal? }

>>>>

\pg
In this section, we describe the work done to make a wide-field image of the entire LOFAR primary beam, centred on EGS. A large-field, image of this field has never yet been made at such low frequencies. Due to technical limitations, the maximum attainable image size (in terms of number of pixels) will limit resolution to about $5''$. As such, the international stations are not used for this part of the project.
A more complete sky model can be created from this image, and %, allowing for one last round of self-calibration on the 3C295 high-resolution model. 
bright out-of-field sources can be subtracted from high-resolution images of the Extended Groth Strip so that their sidelobe emission do not contaminate the final images. To achieve this, we require direction-dependent calibration. Because we do not use the international LOFAR stations, we do not need a high-resolution model of 3C295. %  This ensures that we do not ``hard-set" errors in our high-resolution model caused by other sources in the field.

@@

\subsection{Data \& Observation Properties}
\pg
The dataset used for this PhD project is part of the LOFAR Surveys KSP Tier-1 survey, which consists of a number of 8-hour pointings covering as much of the sky visibile to LOFAR as possible. We analyse one of these pointings, an observation performed on the 28th of August 2014 and centred on the Extended Groth Strip. As part of the Tier-1 survey, it is an 8-hour-long observation. We limit ourselves to analysing only the HBA observation, meaning that we use 365 sub-bands which sample a total bandwidth of 110-182 MHz. We use all Core and Remote LOFAR stations, as well as some of the International stations which were online at the time (specifically the German stations 1-5 and 7, along with the Swedish and the British station).

\pg
The observation is centred on the EGS, and not on 3C295. This means that our calibrator source is not at phase centre, which can potentially introduce problems. Thankfully, new developments in imaging \citepads{2017arXiv171202078T} ensure that the direction-dependent PSF (one of the problems introduced by an off-centre calibrator) is properly modeled and corrected for. This means that direction-independent calibration will be exactly correct in the direction of 3C295, and a good model of this source can be made.

\pg
The data was acquired through the LOFAR Long-Term Archive, and is thus pre-processed and flagged for RFI using the standard tools (NDPPP\citetads{2018ascl.soft04003V} and AOFlagger \citetads{2010ascl.soft10017O}, respectively). The next step is therefore to calibrate the data.

\subsection{Reducing the Data}

@@

\pg
These sources nevertheless remain in the field, as shown in \cref{image.egs.widefield.ddsc.zoom}. Note that the noise level is not constant throughout the image: this is because the array is not uniformly sensitive to the entire sky, but has a specific beam response. Correcting for this beam response in the image-plane gives integrated flux rather than apparent flux. A full catalogue of the sources in this image is given in \cref{appendix.widefield.catalog}, \cref{widefield.survey.catalog} % gives rise to this effect.

@@

\pg
We select 12 sources in the wide-field image, which is shown in \cref{plot.EGS.lofar.widefield}. Our selection is primarily based on whether sources have interesting or peculiar diffuse emission, though some are chosen for being particularly classic examples of radio galaxies. We scan the image from south to north and east to west (scanning upwards then shifting to the right in the image). The source positions are given in \cref{table.egs.sources}.

\begin{figure}[h!]
\centering
\includegraphics[width=0.95\textwidth]{images/{lofar.widefield}.png}
\caption{\label{plot.EGS.lofar.widefield} Full direction-dependent calibrated image of the LOFAR primary beam, centred on the EGS. Image made without international stations.}
\end{figure}

@@

\pg
However, the scientific value of this image is not limited to its use in imaging the Extended Groth Strip using international LOFAR. Much of the image will not be recreated using the international stations, in part due to computational constraints (long run time for such a large task) and in part due to technical limitations (performing direction-dependent calibration using international LOFAR would likely require the use of techniques such as fringe-fitting, which are as-of-yet unimplemented in the DDF-kMS pipeline). As such, we look in the large-field image for interesting or unusual sources, and attempt to match them to known sources in other bands in order to learn about them. 12 sources were selected, and the most interesting were consistently matched to brightest galaxies in clusters. This indicates that there are many sources with very interesting diffuse emission in and around the Extended Groth Strip, and some of which merit further observations to better resolve their turbulent structure.

@@
